---
title: "Visualizing African American achievements throughout the years"
output: 
  html_document:
   theme: flatly
   highlight: zenburn
   number_sections: true
---

# About the data

<i class="fa fa-link"></i><font size="3"> **First.csv**  </font> 

This [data set](https://en.m.wikipedia.org/wiki/List_of_African-American_inventors_and_scientists) has 479 records of African Americans breaking the color barrier across a wide range of
topics. It contains achievements done **first** by African Americans. Breakthroughs if you will.

<i class="fa fa-link"></i><font size="3"> **Science.csv** </font> 

This [data set](https://en.wikipedia.org/wiki/List_of_African-American_firsts) also celebrates the achievements of African Americans, specifically related to Patents and scientific achievements. One of the amazing scientists present in this data set is David Blackwell [cite] an African american mathematician and statistician with significant contributions to game theory, information theory, and Bayesian statistics.

<i class="fa fa-link"></i><font size="3"> **The Black Lives Matter raised fist symbol** </font> 

A clenched raised fist has always been a [symbol of social revolution](https://en.m.wikipedia.org/wiki/Raised_fist#:~:text=As%20early%20as%201914%2C%20a,World%20(IWW)%20in%20197). When the [Black panther group](https://en.m.wikipedia.org/wiki/Black_Panther_Party) was founded in 1966 by [Huey P. Newton](https://en.m.wikipedia.org/wiki/Huey_p._Newton) and [Bobby Seale](https://en.m.wikipedia.org/wiki/Bobby_Seale) to challenge police brutality against the African American community, the black power fist was repeatedly used as an emblem of black liberation. Photos and video footage of the Black Panthers saluting one another with raised fists at conventions, meetings and rallies solidified the symbol as synonymous with the fight for black civil rights.

The black power fist has since been adopted by  several black activists and allies of the black community and has also been used during iconic  events dear to the black community. For example, while receiving their medals during [the 1968 Mexico City Olympics Games](https://en.m.wikipedia.org/wiki/1968_Summer_Olympics),  American sprinters [Tommie Smith](https://en.m.wikipedia.org/wiki/Tommie_Smith) and [John Carlos](https://en.m.wikipedia.org/wiki/John_Carlos) each wore a black glove and raised their fists as the national anthem played. [Nelson Mandela](https://en.m.wikipedia.org/wiki/Nelson_Mandela), the anti-apartheid activists, when released from prison raised a fist alongside his wife. During the [George Floyd protests](https://en.m.wikipedia.org/wiki/George_Floyd_protests), protesters occasionally used it when protesting and spray painted the symbol on buildings and sidewalks.

## Goal for this project

Use data visualization for story telling; to visualize the history of African American achievements.

# Importing the data and the BLM raised fist logo



```{r echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyverse)
library(knitr)
library(stringr)
library(ggplot2)
library(ggthemes)
library(magick)
library(png)
library(grid)
library(scales)
library(readr)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
setwd("~/R- CLASS DATASETS/joanweru.github.io")
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
science<-read_csv("science.csv")
first<-read_csv("first.csv")
black<-readPNG("images-5.png")
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
g<-rasterGrob(black,interpolate = TRUE,width = unit(3,"cm"),x=unit(1,"npc"),hjust=1,vjust=-1,just = "bottom",)
```


# Manipulating and visualizing the data

```{r message=FALSE, warning=FALSE, include=FALSE}
first<-rename(first,"ID"=X1)
```

## `First.csv`

```{r echo=TRUE, message=FALSE, warning=FALSE}
rmarkdown::paged_table(first)

```

We want to visualize the achievements by African-Americans by gender and category. We will use several data wrangling and manipulation functions for that. 

The `gender` column has two levels: female denoted *Female African American Firsts* and male denoted by *African-American firsts* . We want to change them both to simply male and female. We will also count the achievements in each `category` and further look into they're spread across each `gender`.

```{r echo=TRUE, message=FALSE, warning=FALSE,fig.height=5,fig.width=10,fig.align='center',dpi=300}
first%>%group_by(category,gender)%>%count(category,gender)%>%
  ungroup()%>%mutate(category=fct_reorder(category,n))%>%
  mutate(gender=case_when(str_detect(gender,"African-American Firsts")~"Male",TRUE~"Female"))%>%
  ggplot(aes(x=category,y=n,fill=gender))+geom_col(position = "dodge")+
  scale_fill_manual(values = c("salmon3","saddlebrown"))+
  coord_flip()+theme(legend.position = "right",legend.direction = "vertical")+theme_fivethirtyeight()+
  labs(title="MAJOR BREAKTHROUGHS BY AFRICAN AMERICANS BY GENDER AND CATEGORY",caption = "Github: @joanweru")+annotation_custom(g,xmin = -4,ymin=12)+theme(plot.title = element_text(size = 12.5,hjust =0.5))
```


We will also visualize breakthrough achievements as they happened throughout the years.

```{r echo=TRUE, fig.align='center', fig.height=5, fig.width=10, message=FALSE, warning=FALSE, dpi=300}
first%>%mutate(personal=paste(person,accomplishment,sep = "----"))%>%add_count(category)%>%
  ggplot(aes(x=year,y=n,color=category))+geom_point(size=3)+
  scale_color_manual(values = c("chocolate2","sienna4","tomato4","tomato3","salmon4","black","red4","firebrick"))+theme_fivethirtyeight()+theme(legend.position = "right",legend.direction = "vertical",plot.title = element_text(hjust = -0.5,size = 12.5))+
  scale_x_continuous(breaks = seq(1738,2050,30))+annotation_custom(g,xmin = -0.5,ymin = 2)+labs(caption ="Github: @joanweru",title="HISTORY OF MAJOR BREAKTHROUGHS BY AFRICAN AMERICANS IN VARIOUS FIELDS")+coord_cartesian(clip="off")
```


## `Science.csv`

```{r echo=TRUE, message=FALSE, warning=FALSE}
rmarkdown::paged_table(science)

```

```{r message=FALSE, warning=FALSE, include=FALSE}
attach(science)
```


We are interested in showing the achievements done by African American in various field. However looking at the data we have an `occupation_s`column  with entries for various occupations.

```{r echo=TRUE, message=FALSE, warning=FALSE}
str(as.factor(occupation_s))
```

We cannot  visualize all the occupations individually. Our plot will simply be too messy to understand.
We will make a new column `field` where each occupation is grouped into its respective field and we will visualize that instead.

```{r echo=TRUE, message=FALSE, warning=FALSE}

science$field=case_when(str_detect(occupation_s,regex("computer",ignore_case = TRUE))~"Computing",
                         str_detect(occupation_s,regex("engineer|elec|robo|acous",ignore_case = TRUE))~"Engineering",
                         str_detect(occupation_s,regex("math|stat|prob",ignore_case = TRUE))~"Mathematics",
                         str_detect(occupation_s,regex("surgeon|med|nurse|dent|ophtha|psycho|surg|ortho",ignore_case=TRUE))~"Health science",
                         str_detect(occupation_s,regex("astr|atmos",ignore_case = TRUE))~"Space and geography",
                         str_detect(occupation_s,regex("chemi",ignore_case = TRUE))~"Chemistry",
                         str_detect(occupation_s,regex("econ|finan",ignore_case = TRUE))~"Economics",
                         str_detect(occupation_s,regex("physic",ignore_case = TRUE))~"Physics",
                         str_detect(occupation_s,regex("bio|gene|botan|epidem|zoo",ignore_case = TRUE))~"Biology",
                         str_detect(occupation_s,regex("resear",ignore_case = TRUE))~"Research",
                         str_detect(occupation_s,regex("farmer",ignore_case = TRUE))~"Agriculture",
                         str_detect(occupation_s,regex("art|draft|archi",ignore_case = TRUE))~"Design",
                         str_detect(occupation_s,regex("ling|anthro",ignore_case = TRUE))~"Language and Culture",
                         str_detect(occupation_s,regex("inventor|NA",ignore_case = TRUE))~"Inventor",
                        
                      )

```


I noticed two missing values that didn't get matched so I looked up the names of the individuals with the missing values and found that both of them were inventors so i just replaced the *NA* with *Inventor* 

```{r echo=TRUE, message=FALSE, warning=FALSE}
science$field<-science$field%>%replace_na("Inventor")
```

```{r include=FALSE}
attach(science)
```

We now visualize notable inventions and achievements by African-American in various fields throughout each decade.
 
```{r echo=TRUE, message=FALSE, warning=FALSE}
science1<-science%>%mutate(decade=birth-birth%%10)%>%group_by(decade)%>%arrange(field)%>%mutate(index=row_number())%>%ungroup()

science1$decade<-as.numeric(science1$decade)
```




```{r echo=TRUE, fig.align='center', fig.height=5.5, fig.width=15, message=FALSE, warning=FALSE, dpi=300}
science1%>%ggplot(aes(x=decade,y=n,fill=field))+geom_bar(stat = "identity")+theme_fivethirtyeight()+scale_x_continuous(breaks=seq(1730,2010,by=10),labels = paste0(seq(1730,2010,by=10),"s"))+ scale_y_continuous(breaks = seq(0,500,by=30))+
 theme(legend.direction = "vertical",legend.position = "right",plot.title = element_text(size = 15,hjust = 0.5))+
  annotation_custom(g,xmin =max(science1$decade))+coord_cartesian(clip="off")+
  labs(caption = "Github:@joanweru",
       title="HISTORY OF NOTABLE AFRICAN AMERICAN INVENTIONS AND CONTRIBUTIONS IN VARIOUS SCIENTIFIC FIELDS")

```
